---
title: "Analysis of Exponential distribution"
author: "Jose Gonzalez"
date: "6 novembre 2016"
output: html_document
---
#Introduction
In this report we will compare the means and variances of samples taken from a exponential random variable against the theoretical mean and variance of such a distribution.
We will also compare the distribution of the means of the samples to a normal distribution.


#Parameters
```{r setup, include=FALSE}
lambda<-0.2
simulations<-1000
samples<-40
n_breaks<-20 #Number of breaks in histograms
set.seed(42) #Reproducibility of results
```


#Sampled mean vs theoretical mean

##Theoretical mean
Theoretical mean of an exponential random variable is 1/lambda:
```{r}
1/lambda
```

The observed mean of `r prettyNum(simulations,big.mark=',')` drawn from an exponential distribution with lambda `r lambda`:
```{r}
mean(rexp(simulations,lambda))
```

##Sampled mean
Following code will create a distribution of `r simulations` based on the mean of `r samples` values drawn from an exponential distribution:
```{r}
mns = NULL
for (i in 1 : simulations) mns = c(mns, mean(rexp(n=samples,lambda)))
```

The mean of the `r simulations` of `r samples` samples is:
```{r}
mean(mns)
```

Following figure shows the distribution of `r prettyNum(simulations*samples,big.mark=',')` samples drawn from an exponential distribution with lambda `r lambda`
```{r}
rexp_sim<-rexp(simulations*samples,lambda)

hist(rexp_sim,
     breaks=n_breaks,
     main=paste("Histogram of ",simulations*samples," samples from exponential distribution with lambda",lambda),
     xlab="Exponential random variable")

abline(v=1/lambda,lty="dashed",col="red")
text(x=1/lambda,
     y=1000,
     labels=paste("Theoretical mean: ",1/lambda),
     pos=4)
```

The following figure shows the distribution of the mean of `r simulations` distributions of `r samples` drawn from an exponential distribution with lambda `r lambda`
```{r}
hist(mns,
      breaks=n_breaks,
     main=paste("Histogram of means of ",simulations," distributions \nof ",samples, "samples from exponential distribution with lambda",lambda),
     xlab="Mean of the sampled distribution")

abline(v=mean(mns),lty="dashed",col="red")
text(x=mean(mns),
     y=40,
     labels=paste("Observed population mean: ",prettyNum(mean(mns),digits=4)),
     pos=4)
```

##Mean comparion conclusion
The mean of the mean of the sampled distribution `r prettyNum(mean(mns),digits=4)` is close to the theoretical mean 1/lambda: `r prettyNum(1/lambda,digits=4)`

#Theoretical vs sampled variance

##Theoretical variance:
```{r}
1/lambda^2
```

Variance of `r prettyNum(simulations,big.mark=',')` samples:
```{r}
var(rexp(simulations,lambda))
```

##Variance of sampled population
Draw a `r simulations` distributions of `r samples` drawn from an exponential distribution with lambda `r lambda` and take the variance of each of the distributions. Calculate the mean of the variances.
```{r}
var = NULL
for (i in 1 : simulations) var = c(var, var(rexp(n=samples,lambda)))
mean(var)
```

Estimation of population variance based on the mean of the variances of the samples:
```{r}
pvar=(mean(var))*samples/(samples-1)
pvar
```

Following figure shows the distribution of the variances of the sampled distribution:
```{r}
hist(var,
     breaks=n_breaks,
       main=paste("Histogram of variances of ",simulations," distributions \nof ",samples, "samples from exponential distribution with lambda",lambda),
     xlab="Variance")

abline(v=mean(var),lty="dashed",col="red")
text(x=mean(var),
     y=40,
     labels=paste("Observed mean of the variances: ",prettyNum(mean(var),digits=4)),
     pos=4)
```

##Variance comparion conclusion
The estimated variance based on the mean of the variances of the sampled distribution `r prettyNum(pvar,digits=4)` is close to the theoretical mean 1/lambda^2: prettyNum(1/lambda^2,digits=4)

#Comparison to normal distribution
For this comparison, we will calculate the probability of having the 2.5% to 95% percentiles of the means of the sampled distribution inside the 2.5% to 97.5% percentiles of a draw from a normal distribution.

First let's get the quantiles of our distribution of the means
```{r}
quantiles<-c(0.025,.5,.975)
quantile(mns,quantiles)
```

Then get the quantiles of normal a normal distribution with mean equal to observed mean and variance equal to observed variance
```{r}
qnorm(quantiles,mean=mean(mns),sd=sqrt(pvar))
```

Probability of finding the sampled 2.5%-97.5% interval inside normal distribution:
```{r}
prob<-pnorm(q=quantile(mns,quantiles[length(quantiles)]),mean=mean(mns),sd=sqrt(pvar/samples))-pnorm(q=quantile(mns,quantiles[1]),mean=mean(mns),sd=sqrt(pvar/samples))
prob
```

The probability of having drawn the same means from a normal random variable is `r sprintf("%.0f%%",prob*100)`

Following histogram compares the means drawn from the exponential random variable vs the means drawn from a normal distribution
```{r}
mns_norm = NULL
for (i in 1 : simulations) mns_norm = c(mns_norm, mean(rnorm(n=samples,mean=mean(mns),sd=sqrt(pvar))))
mean(var)
hist(mns, col=rgb(0,0,1,1/4),breaks=n_breaks,
     main="Histogram of sampled exponential vs sampled normal",
     xlab="Mean")  # first histogram
hist(mns_norm, col=rgb(1,0,0,1/4), add=T,breaks=n_breaks)  # second
legend("topright", c("Exp", "Normal"), col=c(rgb(0,0,1,1/4), rgb(1,0,0,1/4)), lwd=10)
```

##Comparion to normal distribution conclusion
Both the probability of having the 2.5% to 97.5% percentiles inside the equivalent percentiles of a normal distribuion and the histogram show that the distribution of the means of the samples is normal like.

#Conclusion
This illustrated the Central Limit Theorem. Eventhough the original random variable was exponential, drawing the means from a significant number of samples "normalizes" the distribution.
This is of course only applicable as the number of samples in each distribution and the number of distribuions taken is sufficiently high.