---
title: "Analysis of Exponential distribution"
author: "Jose Gonzalez"
date: "6 novembre 2016"
output: html_document
---

```{r setup, include=FALSE}
lambda<-0.2
simulations<-1000
samples<-40
```


#Distribution of the mean

Theoretical mean of an exponential distribution is 1/lambda:
```{r}
1/lambda
```

The observed mean of `r simulations*samples` drawn from an exponential distribution with lambda `r lambda`:
```{r}
mean(rexp(simulations,lambda))
```

Following code will create a distribution of `r simulations`based on the mean of `r samples` values drawn from an exponential distribution:
```{r}
mns = NULL
for (i in 1 : simulations) mns = c(mns, mean(rexp(n=samples,lambda)))
```

The mean of the `r simulations` of `r samples' samples is:
```{r}
mean(mns)
```

Following figure shows the distribution of `r simulations*samples` samples drawn from an exponential distribution with lambda `r lambda`
```{r}
rexp_sim<-rexp(simulations*samples,lambda)
n_breaks<-20
hist(rexp_sim,
     breaks=n_breaks,
     main=paste("Histogram of ",simulations*samples," samples from exponential distribution with lambda",lambda))

abline(v=1/lambda,lty="dashed",col="red")
text(x=1/lambda,
     y=sum(abs(rexp_sim-1/lambda)<(1/lambda)/n_breaks),
     labels=paste("Theoretical mean: ",1/lambda),
     pos=4)
```

The following figure shows the distribution of the mean of `r simulations` distributions of `r samples` drawn from an exponential distribution with lambda `r lambda`
```{r}
sum(abs(mns-mean(mns))<(max(mns)-min(mns))/n_breaks)
(mean(mns))/n_breaks
```
hist(mns,
      breaks=n_breaks,
     main=paste("Histogram of means of ",simulations," distributions of ",samples, "samples from exponential distribution with lambda",lambda))

abline(v=mean(mns),lty="dashed",col="red")
text(x=mean(mns),
     y=sum(abs(mns-mean(mns))<(mean(mns))/n_breaks),
     labels=paste("Observed population mean: ",mean(mns)),
     pos=4)
```


#Distribution of the variance
```{r}
hist(rexp(simulations,lambda))
```

```{r}
var = NULL
for (i in 1 : simulations) var = c(var, var(rexp(n=samples,lambda)))
hist(var)
```

Theoretical variance deviation:
```{r}
1/lambda^2
```

Variance of `simulations` samples:
```{r}
var(rexp(simulations,lambda))
```

Estimation of population variance based on samples variance:
```{r}
pvar=(mean(var))*samples/(samples-1)
pvar
```


Quantiles of distribution of means
```{r}
quantiles<-c(0.025,.5,.975)
quantile(mns,quantiles)
```

Quantiles of normal distribution
```{r}
qnorm(quantiles,mean=mean(mns),sd=sqrt(pvar/samples))
```

Probability of finding the sampled 2.5%-97.5% interval inside normal distribution:
```{r}
pnorm(q=quantile(mns,quantiles[length(quantiles)]),mean=mean(mns),sd=sqrt(pvar/samples))-pnorm(q=quantile(mns,quantiles[1]),mean=mean(mns),sd=sqrt(pvar/samples))
```

